\chapter*{Acknowledgements}
%\addcontentsline{toc}{chapter}{Acknowledgements}
Extra special thanks go to my supervisor, Rick Cavicchioli, and my co-supervisor, Federico Lauro.
Without their support, useful discussions, editorial assistance and much needed prods along throughout the course of my PhD, I certainly would not have gotten this done or come anywhere near as far.
I am very grateful for the time, energy, patience and hard work that was put into me.
I have learnt a lot!

Thanks to all those involved in the logistics of Antarctic research and especially sample collection.
Expeditioners who collected samples used in this thesis were Rick Cavicchioli, Federico Lauro, Torsten Thomas, Mark Brown, Jeff Hoffman and John Rich.
Thanks to Nico Wanandy for the logistic support at the UNSW end and folks from the Australian Antarctic Division.
Special thanks to John Gibson for contributing his wealth of knowledge about the lakes.

Thanks to the rest of the lab-family from whom I have also learnt a lot and who have been wonderful friends to work with.
Tim Williams, Mark Brown and Matt DeMaere deserve a special mention for their help, advice and support on the projects.
Very special thanks go to David Wilkins and Suhaila Mohd. Omar who have worked alongside with me as fellow PhD companions and without whom I would have been lost.
Thanks to Jodi Richards, Tim Charlton, Sohail Siddiqui and Haluk Ertan for the invaluable lunchtime support group meetings.
Thanks to the past PhD students from the group, Charmaine Ng, Dominic Burg and Davide De Francisci, who left big footprints for me to follow.
Thanks to my extended university family, Jo and Malu, for being my much needed tea and lunch break companions, and Kylie and Pan, for all the mind-expanding biology discussions.

To my at-home science family, thanks for being there as my constant, private support team and my human/non-human pack to run through life with.
Very special thanks to Amy Cain, who has been through the whole gamut of education with me until this point where all the official avenues of higher learning have been exhausted, and to Laura Nolan, who has ended up being as good as a de facto wife as well as thesis sister to me.
Thanks to Julia Suurbach, Meaghan Kukulj, Sissy Reyes, Sophia Slavich and Tim L Williams for being such good friends and inspiring people to be around.
Thanks to Rosko, Mutley, Puggy, Calypso, Icarus, Nakita I and II for the much needed walks, fly-arounds and just hanging around as part of the family.

A very special thanks to my genetic family without whom I definitely would not have been here to do this.
Thanks mum, pop and Jerson, I owe you everything.
Thanks Indy (I know we are only very distantly related really, but it doesn't matter).
