\chapter*{List of Publications}
%\addcontentsline{toc}{chapter}{List of Publications}
Publications and submitted manuscripts arising from my PhD research are listed below.
In all cases, my supervisor Prof Ricardo Cavicchioli and my co-supervisor Dr Federico Lauro were
involved in the research design and editing of the manuscripts.
Where versions of published material, or material submitted for publication appears in this thesis, details of the contributions made by myself and others precede it.

\begin{itemize}

\item \textbf{Sheree Yau}, Federico M. Lauro, Timothy J. Williams, Matthew Z. DeMaere, Mark V. Brown, John Rich, John A.E. Gibson, Ricardo Cavicchioli. 
Strategies of carbon conservation and unusual sulfur biogeochemistry in a hypersaline lake.
\emph{\underline{ISME Journal}}
(submitted), 2013.

\item Khawar S. Siddhiqui, Timothy J. Williams, David Wilkins, \textbf{Sheree Yau}, Michelle A. Allen, Mark V. Brown, Federico M. Lauro, Ricardo Cavicchioli.
Psychrophiles.
\emph{\underline{Annual Review of Earth and Planetary Sciences}}
(in press), 2013.

\item David Wilkins, \textbf{Sheree Yau}, Timothy J. Williams, Michelle Allen, Mark V. Brown, Matthew Z. DeMaere, Federico M. Lauro and Ricardo Cavicchioli.
Key Microbial Drivers in Antarctic Aquatic Environments.
\emph{\underline{FEMS Microbiology Reviews}}
(10.1111/1574-6976.12007), 2012.

\item \textbf{Sheree Yau} and Ricardo Cavicchioli. 
Microbial communities in Antarctic lakes: Entirely new perspectives from metagenomics and metaproteomics. 
\emph{\underline{Microbiology }}
\emph{\underline{Australia}} 32:157--159, 2011.

\item Federico M. Lauro, Matthew Z. DeMaere, \textbf{Sheree Yau}, Mark V. Brown, Charmaine Ng, David Wilkins, Mark J. Raftery, John A.E. Gibson, Cynthia Andrews-Pfannkoch, Matthew Lewis, Jeffery M. Hoffman,Torsten Thomas and Ricardo Cavicchioli. 
An integrative study of a meromictic lake ecosystem in Antarctica. 
\emph{\underline{ISME Journal}}
5:879--895, 2011.

\item \textbf{Sheree Yau}, Federico M. Lauro, Matthew Z. DeMaere, Mark V. Brown, Torsten Thomas, 
Mark J. Raftery, Cynthia Andrews-Pfannkoch, Matthew Lewis, Jeffery M. Hoffman, John A. Gibson and 
Ricardo Cavicchioli. 
Virophage control of antarctic algal host--virus dynamics. 
\emph{\underline{Proceedings of the National Academy of Sciences}}
\emph{\underline{USA}} 108:6163--6168, 2011.

\end{itemize}

